\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-11-05 doc/flexiblesusy-paper.tex
note that CP-violating phases are set to zero
because only real model parameters are currently allowed.
reformulate caption in Table 2
This addresses referee 2, comment 2

The CKM mixing matrix is currently assumed to be diagonal.  The third
generation quark masses are calculated in the \DRbar scheme from the
SLHA user input quantities $m_t^\text{pole}$,
$m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ and

@@

One of \fs's design goals is a short run-time.  In this section we
demonstrate that this goal was achieved by comparing the run-time of
two different sets of CMSSM spectrum generators:
%
\begin{itemize}
\item \emph{Without sfermion flavour violation:} Disallowing sfermion
  flavour violation simplifies the calculation of the pole masses,
  because flavour-off-diagonal sfermion self-energy matrix elements
  don't need to be calculated.  Here we compare \fs's non-flavour
  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
%
\item \emph{With sfermion flavour violation:} Allowing for sfermion
  flavour violation in general increases the run-time of spectrum
  generators, because the full $6\times 6$ sfermion self-energy
  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
  linked against SPheno 3.2.4).  Both spectrum generators are based on
  SARAH's MSSM model file, which allows for sfermion flavour
  violation.
\end{itemize}
%
For the run-time comparison the CKM matrix is set to unity, all CP
violating phases are set to zero and R-parity violation is disabled.
\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort 13.1.3
20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort 13.1.3
20130607.\footnote{Intel's ifort compiler decreases the run-time of
  SPheno and SPhenoMSSM by approximately a factor $1.5$, compared to
  gfortran.}  We're generating $2\cdot 10^{4}$ random CMSSM parameter
points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in \ref{sec:speed-test-slha-template-file}.
The resulting SLHA input file is passed to each spectrum generator and
the (wall-clock) time is measured until the program has finished.  The
average run-times for three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a machine with two
Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators
    for random parameter points.  The first three rows show
    spectrum generators which disallow sfermion flavour violation.  Rows
    $4$--$5$ contain sfermion flavour violating spectrum generators, based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}

>>>>

The CKM mixing matrix is currently set to unity and CP-violating
phases are set to zero.  The third generation quark masses are
calculated in the \DRbar scheme from the SLHA user input quantities
$m_t^\text{pole}$, $m_{b,\text{SM}}^{\text{\MSbar}}(M_Z)$ and

@@

One of \fs's design goals is a short run-time.  In this section we
demonstrate that this goal was achieved by comparing the run-time of
two different sets of CMSSM spectrum generators:
%
\begin{itemize}
\item \emph{Without sfermion flavour violation:} Disallowing sfermion
  flavour violation simplifies the calculation of the pole masses,
  because flavour-off-diagonal sfermion self-energy matrix elements
  don't need to be calculated.  Here we compare \fs's non-flavour
  violating CMSSM spectrum generator FlexibleSUSY-NoFV (version 1.0.0)
  against SPheno (version 3.2.4) and Softsusy (version 3.4.0).
%
\item \emph{With sfermion flavour violation:} Allowing for sfermion
  flavour violation in general increases the run-time of spectrum
  generators, because the full $6\times 6$ sfermion self-energy
  matrices have to be calculated.  Here we compare FlexibleSUSY-FV
  (version 1.0.0) and SPhenoMSSM (generated with SARAH 4.1.0 and
  linked against SPheno 3.2.4).  Both spectrum generators are based on
  SARAH's MSSM model file, which allows for sfermion flavour
  violation.
\end{itemize}
%
For the run-time comparison the CKM matrix is set to unity, all CP
violating phases are set to zero and R-parity violation is disabled.
\fs and Softsusy are compiled with g++ 4.8.0 and Intel ifort 13.1.3
20130607.  SPheno and SPhenoMSSM are compiled with Intel ifort 13.1.3
20130607.\footnote{Intel's ifort compiler decreases the run-time of
  SPheno and SPhenoMSSM by approximately a factor $1.5$, compared to
  gfortran.}  We're generating $2\cdot 10^{4}$ random CMSSM parameter
points with $m_0\in [50,1000]\unit{GeV}$, $m_{1/2}\in
[50,1000]\unit{GeV}$, $\tan\beta\in [1,100]$, $\sign\mu\in \{-1,+1\}$
and $A_0\in [-1000,$ $1000]\unit{GeV}$.  For each point an SLHA input
file is created by appending the values of $m_0$, $m_{1/2}$,
$\tan\beta$, $\sign\mu$, $A_0$ in form of a \code{MINPAR} block to the
SLHA template file given in \ref{sec:speed-test-slha-template-file}.
The resulting SLHA input file is passed to each spectrum generator and
the (wall-clock) time is measured until the program has finished.  The
average run-times for three different CPU types can be found in
\tabref{tab:run-time-comparison}.  The first column shows the run-time
on a Intel Core2 Duo (P8600, $2.40\unit{GHz}$) where only one core was
enabled.  The second column shows the run-time on the same processor
where both cores were enabled.  In the third column a machine with two
Intel Xeon CPUs (L5640, $2.27\unit{GHz}$, $6$ cores) was used.
%
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                            & Intel Core2 Duo    & Intel Core2 Duo   & $2$ $\times$ Intel Xeon\\
                            & (P8600, $1$ core)  & (P8600, $2$ cores)& (L5640, $6$ cores)\\
    \midrule
    FlexibleSUSY-NoFV 1.0.0 & $0.086\unit{s}$    & $0.079\unit{s}$   & $0.060\unit{s}$\\
    SPheno 3.2.4            & $0.119\unit{s}$    & $0.114\unit{s}$   & $0.101\unit{s}$\\
    Softsusy 3.4.0          & $0.175\unit{s}$    & $0.171\unit{s}$   & $0.147\unit{s}$\\
    \midrule
    FlexibleSUSY-FV 1.0.0   & $0.150\unit{s}$    & $0.113\unit{s}$   & $0.074\unit{s}$\\
    SPhenoMSSM 4.1.0        & $0.415\unit{s}$    & $0.401\unit{s}$   & $0.370\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time of CMSSM spectrum generators
    for random parameter points.  The first three rows show three
    spectrum generators with disabled sfermion flavour violation.  Rows
    $4$--$5$ show two spectrum generators with enabled sfermion flavour violation, both based
    on SARAH's MSSM model file.}
  \label{tab:run-time-comparison}
\end{table}
